Verification of concurrent design patterns with data

### 2019-02-25 2018-Arch-pNets/arch-pNets.tex

\begin{abstract}
We provide a solution for the design of safe concurrent systems by
compositional application of verified design patterns\mdash 
called \emph{architectures}\mdash to a small set of functional
components.  To this end, we extend the theory of architectures
developed previously for the BIP framework with the elements necessary
for handling data: definition and operations on data domains, syntax
and semantics of composition operators involving data transfer.  Under
monotonicity assumptions on guards and assignments, composition of
architectures preserves their characteristic safety properties.
To verify that individual architectures do enforce their associated
properties, we provide an encoding into open pNets, an intermediate
model  that supports SMT-based verification.
Additionally, the translation into  open pNet  gives
us a strong insight and a comparison of the coordination mechanisms
provided by the two frameworks.  The approach is illustrated by a case
study based on a previously developed BIP model of a nanosatellite
on-board software.

@@

This article goes one step further in the proof of properties and in the compositionality of architectures, but this step is a significant one: the compositional verification. To prove properties of BIP architectures
it is necessary to have a representation of the BIP architecture in a verifiable format. 
The verification problem has two unbounded parameters: 1) By nature, architectures have holes and are meant to interact with the interfaces of the component that will fill the hole; the properties must hold for all (well-typed) components that can be put inside the hole; 2) BIP interactions can transmit data, and properties might be dependent of the data, the domain of the data is generally huge or unbound and the value of transmitted data might have a significant impact on the properties.
Due to the nature of the problem, we propose to rely on a translation of BIP architectures into pNets, a direct verification of BIP could  be possible but pNets already have verification tools based on a SMT solver, and doing the same work for BIP would require  a large amount of similar development. More interestingly, relying on a translation to pNets allows us to compare the two models.

@@

We propose   to encode BIP architectures
into open pNets and to verify, using SMT techniques,
the properties of the architectures. The encoding also allows us to
compare BIP architectures and  pNet formalism.  
Overall, the encoding allows us to precisely compare the coordination
mechanisms of the two models but also the way component composition is
expressed in the two formalisms. 
%This also gives the opportunity
%to relate two  different visions on compositional verification.

>>>>

\begin{abstract}
We provide a solution for the design of safe concurrent systems by
compositional application of verified design patterns\mdash 
called \emph{architectures}\mdash to a small set of functional
components.  To this end, we extend the theory of architectures
developed previously for the BIP framework with the elements necessary
for handling data: definition and operations on data domains, syntax
and semantics of composition operators involving data transfer.  Under
monotonicity assumptions on guards and assignments, composition of
architectures preserves their characteristic safety properties.
To verify that individual architectures do enforce their associated
properties, we provide an encoding into open pNets, an intermediate
model  that supports SMT-based verification.
Additionally, the translation into  open pNet  \noteEM{ICI}{gives
us a strong insight and a comparison of the coordination mechanisms
provided by the two frameworks}.  The approach is illustrated by a case
study based on a previously developed BIP model of a nanosatellite
on-board software.

@@

This article goes one step further in the proof of properties and in the compositionality of architectures, but this step is a significant one: the compositional verification. To prove properties of BIP architectures
it is necessary to have a representation of the BIP architecture in a verifiable format. 
The verification problem has two unbounded parameters: 1) By nature, architectures have holes and are meant to interact with the interfaces of the component that will fill the hole; the properties must hold for all (well-typed) components that can be put inside the hole; 2) BIP interactions can transmit data, and properties might be dependent of the data, the domain of the data is generally huge or unbound and the value of transmitted data might have a significant impact on the properties.
Due to the nature of the problem, we propose to rely on a translation of BIP architectures into pNets, a direct verification of BIP could  be possible but pNets already have verification tools based on a SMT solver, and doing the same work for BIP would require  a large amount of similar development. More interestingly, relying on a translation to pNets \noteEM{LA}{allows us to compare the two models}.

@@

We propose   to encode BIP architectures
into open pNets and to verify, using SMT techniques,
the properties of the architectures. The encoding also allows us to
compare BIP architectures and  pNet formalism.  
Overall, the encoding \noteEM{ET LA}{allows us to precisely compare the coordination
mechanisms of the two models but also the way component composition is
expressed in the two formalisms}. 
%This also gives the opportunity
%to relate two  different visions on compositional verification.
